---
title: "Multiple linear regression and binary logistic regression models"
author: "Abdellah AitElmouden | Gabriel Abreu |  Jered Ataky | Patrick Maloney"
date: "4/13/2021"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}

library(corrplot)
library(tidyverse)
library(Hmisc)
library(PerformanceAnalytics)
library(mice)
library(gt)
library(caret)
library(bnstruct)
library(VIM)
library(corrr)
library(kableExtra)
library(rpart)
library(gtsummary)

```


## Overview

This assignment is about to explore, to analyze and to model a data set containing approximately 8000
records representing a customer at an auto insurance company. Each record has two response variables. The
first response variable, TARGET_FLAG, is a 1 or a 0. A “1” means that the person was in a car crash. A zero
means that the person was not in a car crash. The second response variable is TARGET_AMT. This value is zero
if the person did not crash their car. But if they did crash their car, this number will be a value greater than zero.
Your objective is to build multiple linear regression and binary logistic regression models on the training data
to predict the probability that a person will crash their car and also the amount of money it will cost if the person
does crash their car. You can only use the variables given to you (or variables that you derive from the variables
provided).

## Data exploration


Below is a short description of the variables of interest in the data set:








```{r include=FALSE}
#Import data 
train_data <- read.csv("https://raw.githubusercontent.com/aaitelmouden/DATA621/master/Homework3/insurance_training_data.csv")


glimpse(train_data)
```


```{r include=FALSE}

# Get rid of INDEX, TARGET_FLAG, TARGET_AMT
train_data1 <- subset(train_data, select = -c(INDEX, TARGET_FLAG, TARGET_AMT))

# Select money variables to clean before summary

new_train_data <- train_data1[, names(train_data1) %in% c("INCOME", "OLDCLAIM", "HOME_VAL", "BLUEBOOK")]

# Clean using string function

new_train_data <- apply(new_train_data, 2, function(y) gsub("\\$","",y))
new_train_data <- apply(new_train_data, 2, function(y) gsub(",","",y))
new_train_data <- apply(new_train_data, 2, as.integer)

# Combine variables

train_data1 <- train_data1[, !names(train_data1) %in% c("INCOME", "OLDCLAIM", "HOME_VAL", "BLUEBOOK")]
train_data1 <- cbind(train_data1, new_train_data)

# Select characters variables to clean before summary

new_train_data <- train_data1[, names(train_data1) %in% c("MSTATUS", "SEX", "EDUCATION", "JOB", "CAR_TYPE", "URBANICITY")]

# Clean using character function

new_train_data <- apply(new_train_data, 2, function(y) gsub("z_","",y))

train_data1 <- train_data1[,!names(train_data1) %in% c("MSTATUS", "SEX", "EDUCATION", "JOB", "CAR_TYPE", "URBANICITY")]
train_data1 <- cbind(train_data1, new_train_data)

```


```{r}

head(train_data1)
```


\newpage

 

\begin{center}
Figure 1 :  Summary Statistics
\end{center}

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Summary table 

train_data1$SEX <- factor(train_data1$SEX)
```

```{r}

train_data1$PARENT1 <- factor(train_data1$PARENT1)
train_data1$CAR_USE <- factor(train_data1$CAR_USE)
train_data1$RED_CAR <- factor(train_data1$RED_CAR)
train_data1$REVOKED <- factor(train_data1$REVOKED)
train_data1$MSTATUS <- factor(train_data1$MSTATUS)
train_data1$EDUCATION <- factor(train_data1$EDUCATION)
train_data1$CAR_TYPE <- factor(train_data1$CAR_TYPE)
train_data1$MSTATUS <- factor(train_data1$MSTATUS)
train_data1$URBANICITY <- factor(train_data1$URBANICITY)
train_data1$JOB <- factor(train_data1$JOB)


```



```{r}

summary(train_data1)
```


\begin{center}
Mean (SD) Median Minimum Maximum 
\end{center}



